A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-07-11 doc/popl2013/intro.tex
intro

A further consequence of placing the same data in multiple logical
regions is that it introduces aliasing which makes it difficult 
to statically figure out which computations can be run in parallel.  
One approach is to outlaw aliasing which allows for a static analysis
to discover all available parallelism and schedule it entirely at
compile-time, avoiding any runtime overhead.  Languages such as
Sequoia and DPJ take this approach and allow only a single static partition of any region 
to eliminate aliasing at the cost of the flexibility of their language.

@@

The important difference between Legion and the more static approaches
is that Legion allows aliasing, but does so in a structured way that minimizes
runtime overheads.  The key observation is that
static alias analysis is hard, but dynamic analysis is very easy
and relatively cheap
when done at the granularity of regions instead of individual heap
locations.  Thus, Legion falls between fully static systems, such as
Sequoia and DPJ, and fully dynamic approaches such as transactional
memory.  Legion still has a significant static component; in
particular, the required privileges for function calls are checked statically
and region
pointer dereferences are checked statically.  The dynamic alias tests
for independence are done at the granularity of regions and one check 
on a region
often serves to prove the independence of many subsequent uses of that region.
In contrast, because transactional memory has no mechanism
for grouping heap data into larger collections, it must test the
overlap between two sets of data on each individual memory access, which is
significantly more expensive.

>>>>

A further consequence of placing the same data in multiple logical
regions is that it introduces aliasing which makes it difficult 
to statically determine which computations can be run in parallel.  
One approach is to outlaw aliasing which allows for a static analysis
to discover all available parallelism and schedule it entirely at
compile-time, avoiding any runtime overhead.  Languages such as
Sequoia and DPJ take this approach and allow only a single static partition of any region 
to eliminate region aliasing, at some cost in expressiveness.

@@

The important difference between Legion and the more static approaches
is that Legion allows region aliasing, but does so in a structured way that minimizes
runtime overheads.  The key observation is that
static alias analysis is hard, but dynamic analysis is very easy
and relatively cheap
when done at the granularity of regions instead of individual heap
locations.  Thus, Legion falls between fully static systems, such as
Sequoia and DPJ, and fully dynamic approaches such as transactional
memory.  Legion still has a significant static component; in
particular, the required privileges for function calls
and region
pointer dereferences are checked statically.  The dynamic alias checks
for independence are done at the granularity of regions and one check 
on a region
often serves to prove the independence of many subsequent uses of that region.
In contrast, because transactional memory has no mechanism
for grouping heap data into larger collections, it must test the
overlap between two sets of data on each individual memory access, which is
significantly more expensive.
